Parsing Postfix Log Files

### 2009-04-28 doc/logparser-architecture.tex
Changes made after final checks.

    % Reduce inter-item spacing because each item is a single line.
    \squeezeitems{}

@@

Excessive CPU time is usually consumed when a regex fails to match an
input, and the regex engine backtracks many times because of alteration or
nested quantifiers; successful matching is generally quite fast with such
regexes, so problematic regexes may go unnoticed for a long time.  For
example, with most regex engines matching double quoted strings with
\verb!\"([^"\\]+|\\.)*"\!  is very fast when there is a match, but when the
match fails its computational complexity for a string of length $n$ is
$O(2^{n})$; see~\cite{mastering-regular-expressions} for in-depth
discussion of nested quantifiers, backtracking, alteration, and capturing
groups.  Pathological regexes that consume excessive CPU time can be
difficult to detect, whether by visual inspection or by machine inspection,
but if a regex is converted to a \acronym{FA} or the internal
representation used by the regex engine, it may be possible to determine if
nested quantifiers or other troublesome constructs are present.  Modern
regex engines have addressed many of these problems, e.g.\ the regex to
match double quoted strings given above fails immediately with Perl 5.10,
regardless of the input length, because the regex engine looks for both of
the required double quotes first.  Similarly, Perl 5.10's regex engine
optimises alterations starting with literal text into a trie, which has
matching time proportional to the length of the alternatives, rather than
the number of alternatives.  Perl regexes can use \verb!(?>pattern)!, which
matches \verb!pattern! the first time the regex engine passes over it, and
does not change what it originally matched if the regex engine backtracks
over it, alleviating problems caused by excessive backtracking; Prolog
users will notice a similarity to the \verb'!' (cut) operator.  A
presentation showing some of Perl 5.10's new regex features is available at
\urlLastChecked{http://www.regex-engineer.org/slides/perl510_regex.html}{2009/03/03}.

>>>>

@@

Excessive CPU time is usually consumed when a regex fails to match an
input, and the regex engine backtracks many times because of alteration or
nested quantifiers; successful matching is generally quite fast with such
regexes, so problematic regexes may go unnoticed for a long time.  For
example, with most regex engines matching double quoted strings with
\verb!\"([^"\\]+|\\.)*"\!  is very fast when a match can be found, but when
the match fails its computational complexity for a string of length $n$ is
$O(2^{n})$; see~\cite{mastering-regular-expressions} for in-depth
discussion of nested quantifiers, backtracking, alteration, and capturing
groups.  Pathological regexes that consume excessive CPU time can be
difficult to detect, whether by visual inspection or by machine inspection,
but if a regex is converted to a \acronym{FA} or the internal
representation used by the regex engine, it may be possible to determine if
nested quantifiers or other troublesome constructs are present.  Modern
regex engines have addressed many of these problems, e.g.\ the regex to
match double quoted strings given above fails immediately with Perl 5.10,
regardless of the input length, because the regex engine looks for both of
the required double quotes first.  Similarly, Perl 5.10's regex engine
optimises alterations starting with literal text into a trie, which has
matching time proportional to the length of the alternatives, rather than
the number of alternatives.  Perl regexes can use \verb!(?>pattern)!, which
matches \verb!pattern! the first time the regex engine passes over it, and
does not change what it originally matched if the regex engine backtracks
over it, alleviating problems caused by excessive backtracking; Prolog
users will notice a similarity to the \verb'!' (cut) operator.  A
presentation showing some of Perl 5.10's new regex features is available at
\urlLastChecked{http://www.regex-engineer.org/slides/perl510_regex.html}{2009/03/03}.

### 2009-04-28 doc/logparser-background.tex
Changes made after final checks.

    \item Checking if the client satisfies a domain's \acronym{SPF}
        records, \urlLastChecked{http://www.openspf.org/}{2009/04/21}.
        \acronym{SPF} records specify which mail servers are allowed to
        send mail using sender addresses from a particular domain.  The
        intention is to reduce spam from faked sender addresses,
        backscatter, and joe~jobs~\glsadd{Joe-job}.  There has been
        considerable resistance to \acronym{SPF} because it breaks or
        vastly complicates some commonly-used features of \acronym{SMTP},
        e.g.\ forwarding mail from one domain to another when a user moves.

>>>>

    \item Checking if the client satisfies a domain's \acronym{SPF}
        records, \urlLastChecked{http://www.openspf.org/}{2009/04/21}.
        \acronym{SPF} records specify which mail servers are allowed to
        send mail using sender addresses from a particular domain.  The
        intention is to reduce spam from faked sender addresses,
        backscatter\glsadd{Backscatter}, and joe~jobs\glsadd{Joe-job}.
        There has been considerable resistance to \acronym{SPF} because it
        breaks or vastly complicates some commonly-used features of
        \acronym{SMTP}, e.g.\ forwarding mail from one domain to another
        when a user moves.

### 2009-04-28 doc/logparser-implementation.tex
Changes made after final checks.

\parsername{} deals with all the eccentricities and oddities of parsing
Postfix log files, presenting the resulting data in a normalised, simple to
use representation.  Unfortunately, dealing with the complications that
arise sometimes requires the parser to discard log lines (e.g.\
\sectionref{aborted delivery attempts}), and, less frequently, to discard a
data structure (e.g.\ \sectionref{timeouts during data phase}).
\parsername{} can parse log files from Postfix version 2.2 through to
version 2.5, and should be able to parse log files from later versions with
only minor modifications or an updated ruleset.  The parser's task is to
follow the journey each mail takes through Postfix, combining the data
captured by rules into a coherent whole, and saving it in a useful and
consistent form.  The intermingling of
log lines from different mails immediately rules out the possibility of
handling each mail in isolation; the parser must handle multiple mails in
parallel, each potentially at a different stage in its journey, without any
interference between mails --- except in the minority of cases where
intra-mail interference is required, e.g.\ mail re-injected for forwarding
(\sectionref{Re-injected mails}).  The best way to deal with intermingling
of log lines is to maintain state information for every unfinished mail,
and to manipulate the appropriate mail correctly for each log line
encountered.

@@

    \item The log files are whole and complete: nothing has been removed,
        either deliberately or accidentally (e.g.\ log file rotation gone
        awry, file system filling up, logging system unable to cope with
        the volume of log messages).  On a well run mail server it is
        extremely unlikely that any of these problems will arise, though
        the likelihood increases when suffering from a deluge of spam or a
        mail loop.  When parsing individual log files in isolation, it is
        highly likely that some mails will have log lines in previous log
        files, and others will have log lines in subsequent log files; to
        alleviate this problem \parsername{} supports saving and loading
        its state tables, so they can be preserved between log files.

@@

            \item [timestamp] The time the line was logged at.

@@

            \item [pid] The \acronym{pid} of the process that logged the
                line.

            \item [host] The hostname of the server the line was logged on.

@@

    \item [SMTPD\_WATCHDOG] \daemon{smtpd} processes have a watchdog timer
        to deal with unusual situations; after five hours the timer will
        expire and the \daemon{smtpd} will exit.  This occurs very
        infrequently, because there are many other timeouts that should
        occur in the intervening hours, e.g.\ timeouts for DNS requests or
        timeouts reading data from the client.  The active connection for
        that \daemon{smtpd} must be cleaned up, entered in the database,
        and deleted from the state tables.

@@

    \item [Pre-process the file.]  The modified algorithm begins by
        leveraging the knowledge gained from writing rules and developing
        \parsername{}; it performs many substitutions on the input log
        lines, replacing commonly occurring variable terms (e.g.\ email
        addresses, IP addresses, the standard start of rejection messages)
        with keywords described in \sectionref{regex components}.  The
        purpose of this step is to utilise existing knowledge to create
        more accurate regexes; it replaces many variable words with fixed
        words, improving the subsequent classification of words as fixed or
        variable.  Regex metacharacters in the log line will be escaped, to
        avoid generating invalid or incorrect regexes.  The altered log
        lines are written to a temporary file, which the next stage will
        use instead of the original input file.

>>>>

\parsername{} deals with all the eccentricities and oddities of parsing
Postfix log files, presenting the resulting data in a normalised, simple to
use representation.  Unfortunately, dealing with the complications that
arise sometimes requires the parser to discard log lines (e.g.\
\sectionref{aborted delivery attempts}), and, less frequently, to discard a
data structure (e.g.\ \sectionref{timeouts during data phase}).
\parsername{} can parse log files from Postfix version 2.2 through to
version 2.5, and should parse log files from later versions with only minor
modifications or an updated ruleset.  The parser's task is to follow the
journey each mail takes through Postfix, combining the data captured by
rules into a coherent whole, and saving it in a useful and consistent form.
The intermingling of log lines from different mails immediately rules out
the possibility of handling each mail in isolation; the parser must handle
multiple mails in parallel, each potentially at a different stage in its
journey, without any interference between mails --- except in the minority
of cases where intra-mail interference is required, e.g.\ mail re-injected
for forwarding (\sectionref{Re-injected mails}).  The best way to deal with
intermingling of log lines is to maintain state information for every
unfinished mail, and to manipulate the appropriate mail correctly for each
log line encountered.

@@

    \item The log files are whole and complete: nothing has been removed,
        either deliberately or accidentally (e.g.\ log file rotation gone
        awry, filesystem filling up, logging system unable to cope with the
        volume of log messages).  On a well run mail server it is extremely
        unlikely that any of these problems will arise, though the
        likelihood increases when suffering from a deluge of spam or a mail
        loop.  When parsing individual log files in isolation, it is highly
        likely that some mails will have log lines in previous log files,
        and others will have log lines in subsequent log files; to
        alleviate this problem \parsername{} supports saving and loading
        its state tables, so they can be preserved between log files.

@@

            \item [timestamp] The time the log line was written to the log
                file.

@@

            \item [pid] The \acronym{pid} of the process that generated the
                log line.

            \item [host] The hostname of the server the log line was
                generated on.

@@

    \item [SMTPD\_WATCHDOG] \daemon{smtpd} processes have a watchdog timer
        to deal with unusual situations; after five hours the timer will
        expire and the \daemon{smtpd} will exit.  This occurs very
        infrequently, because many other timeouts should occur in the
        intervening hours, e.g.\ timeouts for DNS requests or timeouts
        reading data from the client.  The active connection for that
        \daemon{smtpd} must be cleaned up, entered in the database, and
        deleted from the state tables.

@@

    \item [Pre-process the input.]  The modified algorithm begins by
        leveraging the knowledge gained from writing rules and developing
        \parsername{}; it performs many substitutions on the input log
        lines, replacing commonly occurring variable terms (e.g.\ email
        addresses, IP addresses, the standard start of rejection messages)
        with keywords described in \sectionref{regex components}.  The
        purpose of this step is to utilise existing knowledge to create
        more accurate regexes; it replaces many variable words with fixed
        words, improving the subsequent classification of words as fixed or
        variable.  Regex metacharacters in the log line will be escaped, to
        avoid generating invalid or incorrect regexes.  The altered log
        lines are written to a temporary file, which the next stage will
        use instead of the original input file.

### 2009-04-28 doc/logparser-other-parsers-review.tex
Changes made after final checks.

Similar to AWStats and Logrep, Lire attempts to correlate log lines by
queueid, but not by \acronym{pid}, so the complete list of recipients for
each delivered mail should be available; it does not attempt to deal with
the other complications described in \sectionref{complications}.  When
testing Lire on the \numberOFlogFILES{} test log files it performed
reasonably well: the numbers it reports appear accurate, and the subset
verified by the author were correct.  Its report provides summaries of:

@@

>>>>

Similar to AWStats and Logrep (\sectionref{Logrep review}), Lire attempts
to correlate log lines by queueid, but not by \acronym{pid}, so the
complete list of recipients for each delivered mail should be available; it
does not attempt to deal with the other complications described in
\sectionref{complications}.  When testing Lire on the \numberOFlogFILES{}
test log files it performed reasonably well: the numbers it reports appear
accurate, and the subset verified by the author were correct.  Its report
provides summaries of:

@@

\label{Logrep review}

### 2009-04-28 doc/logparser-results.tex
Changes made after final checks.

The second criterion the parser is evaluated on is its coverage of Postfix
log files.  There are two types of coverage to be evaluated: what
proportion of log lines are correctly recognised by the ruleset, and what
proportion of mail delivery attempts are correctly understood and
reconstructed by the actions?  The former is a requirement for the latter
to be achieved, and the latter is important because the data provided by
\parsername{} must be both complete and correct for it to be of use to
others.

>>>>

The second criterion the parser is evaluated on is its coverage of Postfix
log files.  Two kinds of coverage need to be evaluated: what proportion of
log lines are correctly recognised by the ruleset, and what proportion of
mail delivery attempts are correctly understood and reconstructed by the
actions?  The former is a requirement for the latter to be achieved, and
the latter is important because the data provided by \parsername{} must be
both complete and correct for it to be of use to others.
